---
title: "p8105_hw3_cr3442"
author: "Cheng Rao"
date: "2024-10-15"
output: html_document
---

title: "p8105_hw3_cr3442"
author: "Cheng Rao"
date: "2024-10-15"
output: github_document
---

---
title: "p8105_hw3_cr3442"
author: "Cheng Rao"
date: "2024-10-15"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First of all, let me explain why I did not submit multiple versions of the assignment and did not reflect my modification process. It is like this. I had completed part of my assignment before this week, but did not commit or push. Last Saturday, my original computer broke down. I tried to fix the computer during the weekend but failed. For this reason, I used my classmate's computer to complete this week's homework, so I did not submit the previous modification process (because I can't use icloud). And I feel that I can't submit my code through the push method that comes with R studio, but need to submit it by manually adding it on github. I am very sorry for the trouble caused to you in correcting your homework.

# Problem 2

```{r}
library(tidyverse)
library(readr)
```

To solve this problem, first let us load the two required datasets, and check them to make sure they are OK.

```{r load}

getwd()
demographics <- read_csv("/Users/raocheng/Desktop/p8105_hw3_cr3442/data/demographics.csv")
accelerometer <- read_csv("/Users/raocheng/Desktop/p8105_hw3_cr3442/data/accelerometer.csv")

# Preview the first few rows of each dataset
head(demographics)
head(accelerometer)
```

Because the demographics data does not meet the requirements for subsequent operations, we perform some cleaning.

```{r clean}

# Omit the first four rows of data because of the column of age
cleaned_demographics <- read_csv("/Users/raocheng/Desktop/p8105_hw3_cr3442/data/demographics.csv", skip = 4)

head(cleaned_demographics)
str(cleaned_demographics)

```

Next, it is time to merge these data.

```{r merge}

merged_data <- inner_join(cleaned_demographics, accelerometer, by = "SEQN")

# Check the merged data
head(merged_data)
colnames(merged_data)

```
Next, let’s summarize the data.

```{r summary}

# Count the number of people by education level and gender
summary_table <- merged_data %>%
  group_by(education, sex) %>%
  summarize(count = n(), .groups = 'drop')


print(summary_table)
```

Then visualize the data.

```{r visualize}

# Install and load the ggplot2 package
if (!require("ggplot2")) install.packages("ggplot2")
library(ggplot2)

# Draw the age distribution graph
ggplot(merged_data, aes(x = age, fill = as.factor(sex))) +
  geom_histogram(bins = 30, alpha = 0.6, position = "identity") +
  facet_wrap(~ education) +
  labs(title = "Age Distribution by Gender and Education Level",
       x = "Age", y = "Count", fill = "Sex") +
  theme_minimal()
```

The accelerometer data contains minute-by-minute activity data for 24 hours, and we need to summarize the activity for each participant.

```{r total_act}

# Summarize the total activity for each participant
total_activity <- merged_data %>%
  rowwise() %>%
  mutate(total_activity = sum(c_across(starts_with("MIMS")), na.rm = TRUE)) %>%
  ungroup()


head(total_activity)
```

Then we will plot a scatter plot of each participant's total activity versus age, distinguishing between gender and education level.

```{r ggplot}

# Plot the relationship between total activity and age
ggplot(total_activity, aes(x = age, y = total_activity, color = as.factor(sex))) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "loess", se = FALSE) +
  facet_wrap(~ education) +
  labs(title = "Total Activity vs Age by Gender and Education Level",
       x = "Age", y = "Total Activity", color = "Sex") +
  theme_minimal()
```

Finally, create a three-panel plot showing the 24-hour activity time trajectory for each education level, color-coded by gender.

```{r covert}

# Convert activity data for each minute within 24 hours to long format
activity_long <- merged_data %>%
  pivot_longer(cols = starts_with("min"), names_to = "minute", values_to = "activity") %>%
  mutate(minute = as.numeric(gsub("min", "", minute)))

# Draw 24-hour activity trajectory
ggplot(activity_long, aes(x = minute, y = activity, color = sex)) +
  geom_smooth(se = FALSE) +
  facet_wrap(~ education) +
  labs(title = "24-hour activity time trajectory by education level",
       x = "Minutes of the Day", y = "Activity Level", color = "gender") +
  theme_minimal()

```


# Problem 3

```{r}
library(tidyverse)
library(lubridate)
```

```{r}
setwd("/Users/raocheng/Desktop/p8105_hw3_cr3442")
getwd()

Jan_2020_Citi <- read_csv("/Users/raocheng/Desktop/p8105_hw3_cr3442/data/Jan_2020_Citi.csv")
Jan_2024_Citi <- read_csv("/Users/raocheng/Desktop/p8105_hw3_cr3442/data/Jan_2024_Citi.csv")
July_2020_Citi <- read_csv("/Users/raocheng/Desktop/p8105_hw3_cr3442/data/July_2020_Citi.csv")
July_2024_Citi <- read_csv("/Users/raocheng/Desktop/p8105_hw3_cr3442/data/July_2024_Citi.csv")

# Preview the first few rows of each dataset
head(Jan_2020_Citi)
head(Jan_2024_Citi)
head(July_2020_Citi)
head(July_2024_Citi)
```
```{r}
# Merge all datasets
citi_bike_data <- bind_rows(
  Jan_2020_Citi %>% mutate(year = 2020, month = "January"),
  July_2020_Citi %>% mutate(year = 2020, month = "July"),
  Jan_2024_Citi %>% mutate(year = 2024, month = "January"),
  July_2024_Citi %>% mutate(year = 2024, month = "July")
)

head(citi_bike_data)
```

```{r}
# Clean and organize data
citi_bike_data_cleaned <- citi_bike_data %>%
  mutate(day_of_week = weekdays,  # Use existing weekday column
         duration_minutes = duration,  
         member_casual = ifelse(member_casual == "member", "Member", "Casual")) %>%
  drop_na()  

# check the cleaned data
head(citi_bike_data_cleaned)
```

```{r}

ride_summary <- citi_bike_data_cleaned %>%
  group_by(year, month, member_casual) %>%
  summarize(total_rides = n(), .groups = 'drop')

# Generate tabular output
ride_summary_table <- ride_summary %>%
  pivot_wider(names_from = member_casual, values_from = total_rides, values_fill = 0)

print(ride_summary_table)
```

```{r}
# Filter out data from July 2024
july_2024_data <- citi_bike_data_cleaned %>%
  filter(year == 2024, month == "July")

# Find the most popular starting stations
top_start_stations <- july_2024_data %>%
  group_by(start_station_name) %>%
  summarize(total_rides = n(), .groups = 'drop') %>%
  arrange(desc(total_rides)) %>%
  slice_head(n = 5)

print(top_start_stations)
```

```{r}
# Calculate the median riding time for each group
median_ride_duration <- citi_bike_data_cleaned %>%
  group_by(year, month, day_of_week) %>%
  summarize(median_duration = median(duration_minutes), .groups = 'drop')

# Draw a chart
ggplot(median_ride_duration, aes(x = day_of_week, y = median_duration, fill = month)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ year) +
  labs(title = "Median cycling duration by week, month, and year",
       x = "week", y = "Median riding time (minutes)", fill = "month") +
  theme_minimal()
```

```{r}
# Filter data of 2024
data_2024 <- citi_bike_data_cleaned %>%
  filter(year == 2024)

# Plotting the distribution of riding time
ggplot(data_2024, aes(x = duration_minutes, fill = interaction(member_casual, rideable_type))) +
  geom_density(alpha = 0.6) +
  facet_wrap(~ month) +
  labs(title = "Distribution of riding time in 2024 (by month, membership status and vehicle type)",
       x = "Cycling time (minutes)", y = "density", fill = "membership status and vehicle type") +
  theme_minimal() +
  xlim(0, 240) 
```